---
title: "MGMT 3345 Unit 2 Project"
author: "Brian Knowlton"
output: html_notebook
---

Run the following R Chunk to ensure you have enough decimal places for rounding and that values are not in scientific notation.
```{r}
options(digits=10)
options(scipen=999)
```

Load the packages and the BIL data set (1st worksheet of BIL.xlsx).
```{r}
# install.packages(c("readxl", "epiDisplay")) # run once if not already installed
library(readxl)
library(epiDisplay)
BIL <- read_excel("data/BIL.xlsx", sheet = "BIL")
dim(BIL)
```



**Descriptive Statistics**
Load epiDisplay package for the tab1 function.
```{r}
tab1(BIL$GenderCategorical, graph = FALSE)
tab1(BIL$RaceCategorical, graph = FALSE)
tab1(BIL$EthnicityCategorical, graph = FALSE)
tab1(BIL$Educ, graph = FALSE)  # 1 = less than HS ... 6 = doctorate (keeps the levels in order)
tab1(BIL$Marital, graph = FALSE)  # 1 = married/cohabitating, 2 = single, 8 = prefer not to say
tab1(BIL$Utah, graph = FALSE)
```



Average salary (Income_1, in $1000's)
```{r}
mean(BIL$Income_1, na.rm=TRUE)
sd(BIL$Income_1, na.rm=TRUE)
```



Female/Male (Gender)
**Utah/NotUtah (Utah)**  <- my comparison group (last name F through P)
White/nonWhite (White)

Note: R orders the groups alphabetically, so in every t.test and var.test below group 1 = NotUtah and group 2 = Utah.
t = (mean NotUtah - mean Utah) / SE and F = var(NotUtah) / var(Utah).
All tests are one-tailed. The alternative is set in the direction of the observed difference:
"greater" tests NotUtah > Utah (or var NotUtah > var Utah) and "less" tests NotUtah < Utah.


**Differences Not Utah/Utah Intrinsic Rewards (Utah)**

Interesting Job (JobCharacteristics_4)
```{r}
#mean of whole sample
mean(BIL$JobCharacteristics_4, na.rm=TRUE)

#standard deviation of whole sample
sd(BIL$JobCharacteristics_4, na.rm=TRUE)

#mean for NotUtah and Utah subsets
tapply(BIL$JobCharacteristics_4, BIL$Utah, mean, na.rm=TRUE)

#differences of means test (one-tailed)
t.test(BIL$JobCharacteristics_4~BIL$Utah, mu = 0, alternative = "less")

#standard deviation for NotUtah and Utah subsets
tapply(BIL$JobCharacteristics_4, BIL$Utah, sd, na.rm=TRUE)

#differences of variation test (one-tailed)
var.test(BIL$JobCharacteristics_4~BIL$Utah, ratio = 1, alternative = "greater")
```

Job autonomy (JobCharacteristics_5)
```{r}
#mean of whole sample
mean(BIL$JobCharacteristics_5, na.rm=TRUE)

#standard deviation of whole sample
sd(BIL$JobCharacteristics_5, na.rm=TRUE)

#mean for NotUtah and Utah subsets
tapply(BIL$JobCharacteristics_5, BIL$Utah, mean, na.rm=TRUE)

#differences of means test (one-tailed)
t.test(BIL$JobCharacteristics_5~BIL$Utah, mu = 0, alternative = "less")

#standard deviation for NotUtah and Utah subsets
tapply(BIL$JobCharacteristics_5, BIL$Utah, sd, na.rm=TRUE)

#differences of variation test (one-tailed)
var.test(BIL$JobCharacteristics_5~BIL$Utah, ratio = 1, alternative = "greater")
```

Help Others (JobCharacteristics_6)
```{r}
#mean of whole sample
mean(BIL$JobCharacteristics_6, na.rm=TRUE)

#standard deviation of whole sample
sd(BIL$JobCharacteristics_6, na.rm=TRUE)

#mean for NotUtah and Utah subsets
tapply(BIL$JobCharacteristics_6, BIL$Utah, mean, na.rm=TRUE)

#differences of means test (one-tailed)
t.test(BIL$JobCharacteristics_6~BIL$Utah, mu = 0, alternative = "less")

#standard deviation for NotUtah and Utah subsets
tapply(BIL$JobCharacteristics_6, BIL$Utah, sd, na.rm=TRUE)

#differences of variation test (one-tailed)
var.test(BIL$JobCharacteristics_6~BIL$Utah, ratio = 1, alternative = "greater")
```

Job useful to society (JobCharacteristics_7)
```{r}
#mean of whole sample
mean(BIL$JobCharacteristics_7, na.rm=TRUE)

#standard deviation of whole sample
sd(BIL$JobCharacteristics_7, na.rm=TRUE)

#mean for NotUtah and Utah subsets
tapply(BIL$JobCharacteristics_7, BIL$Utah, mean, na.rm=TRUE)

#differences of means test (one-tailed)
t.test(BIL$JobCharacteristics_7~BIL$Utah, mu = 0, alternative = "less")

#standard deviation for NotUtah and Utah subsets
tapply(BIL$JobCharacteristics_7, BIL$Utah, sd, na.rm=TRUE)

#differences of variation test (one-tailed)
var.test(BIL$JobCharacteristics_7~BIL$Utah, ratio = 1, alternative = "less")
```

Job satisfaction (JobSat)
```{r}
#mean of whole sample
mean(BIL$JobSat, na.rm=TRUE)

#standard deviation of whole sample
sd(BIL$JobSat, na.rm=TRUE)

#mean for NotUtah and Utah subsets
tapply(BIL$JobSat, BIL$Utah, mean, na.rm=TRUE)

#differences of means test (one-tailed)
t.test(BIL$JobSat~BIL$Utah, mu = 0, alternative = "less")

#standard deviation for NotUtah and Utah subsets
tapply(BIL$JobSat, BIL$Utah, sd, na.rm=TRUE)

#differences of variation test (one-tailed)
var.test(BIL$JobSat~BIL$Utah, ratio = 1, alternative = "greater")
```


**Differences Not Utah/Utah Extrinsic Rewards (Utah)**

Income in $1000's (Income_1)
```{r}
#mean of whole sample
mean(BIL$Income_1, na.rm=TRUE)

#standard deviation of whole sample
sd(BIL$Income_1, na.rm=TRUE)

#mean for NotUtah and Utah subsets
tapply(BIL$Income_1, BIL$Utah, mean, na.rm=TRUE)

#differences of means test (one-tailed)
t.test(BIL$Income_1~BIL$Utah, mu = 0, alternative = "greater")

#standard deviation for NotUtah and Utah subsets
tapply(BIL$Income_1, BIL$Utah, sd, na.rm=TRUE)

#differences of variation test (one-tailed)
var.test(BIL$Income_1~BIL$Utah, ratio = 1, alternative = "greater")
```

Perception of Income (JobCharacteristics_2)
```{r}
#mean of whole sample
mean(BIL$JobCharacteristics_2, na.rm=TRUE)

#standard deviation of whole sample
sd(BIL$JobCharacteristics_2, na.rm=TRUE)

#mean for NotUtah and Utah subsets
tapply(BIL$JobCharacteristics_2, BIL$Utah, mean, na.rm=TRUE)

#differences of means test (one-tailed)
t.test(BIL$JobCharacteristics_2~BIL$Utah, mu = 0, alternative = "greater")

#standard deviation for NotUtah and Utah subsets
tapply(BIL$JobCharacteristics_2, BIL$Utah, sd, na.rm=TRUE)

#differences of variation test (one-tailed)
var.test(BIL$JobCharacteristics_2~BIL$Utah, ratio = 1, alternative = "greater")
```

Job security (JobCharacteristics_1)
```{r}
#mean of whole sample
mean(BIL$JobCharacteristics_1, na.rm=TRUE)

#standard deviation of whole sample
sd(BIL$JobCharacteristics_1, na.rm=TRUE)

#mean for NotUtah and Utah subsets
tapply(BIL$JobCharacteristics_1, BIL$Utah, mean, na.rm=TRUE)

#differences of means test (one-tailed)
t.test(BIL$JobCharacteristics_1~BIL$Utah, mu = 0, alternative = "less")

#standard deviation for NotUtah and Utah subsets
tapply(BIL$JobCharacteristics_1, BIL$Utah, sd, na.rm=TRUE)

#differences of variation test (one-tailed)
var.test(BIL$JobCharacteristics_1~BIL$Utah, ratio = 1, alternative = "greater")
```

Promotional opportunities (JobCharacteristics_3)
```{r}
#mean of whole sample
mean(BIL$JobCharacteristics_3, na.rm=TRUE)

#standard deviation of whole sample
sd(BIL$JobCharacteristics_3, na.rm=TRUE)

#mean for NotUtah and Utah subsets
tapply(BIL$JobCharacteristics_3, BIL$Utah, mean, na.rm=TRUE)

#differences of means test (one-tailed)
t.test(BIL$JobCharacteristics_3~BIL$Utah, mu = 0, alternative = "greater")

#standard deviation for NotUtah and Utah subsets
tapply(BIL$JobCharacteristics_3, BIL$Utah, sd, na.rm=TRUE)

#differences of variation test (one-tailed)
var.test(BIL$JobCharacteristics_3~BIL$Utah, ratio = 1, alternative = "greater")
```


**Differences Not Utah/Utah Work Quality (Utah)**

Job engagement (Eng_1)
```{r}
#mean of whole sample
mean(BIL$Eng_1, na.rm=TRUE)

#standard deviation of whole sample
sd(BIL$Eng_1, na.rm=TRUE)

#mean for NotUtah and Utah subsets
tapply(BIL$Eng_1, BIL$Utah, mean, na.rm=TRUE)

#differences of means test (one-tailed)
t.test(BIL$Eng_1~BIL$Utah, mu = 0, alternative = "less")

#standard deviation for NotUtah and Utah subsets
tapply(BIL$Eng_1, BIL$Utah, sd, na.rm=TRUE)

#differences of variation test (one-tailed)
var.test(BIL$Eng_1~BIL$Utah, ratio = 1, alternative = "greater")
```

Meaningful career (MW_PM_1)
```{r}
#mean of whole sample
mean(BIL$MW_PM_1, na.rm=TRUE)

#standard deviation of whole sample
sd(BIL$MW_PM_1, na.rm=TRUE)

#mean for NotUtah and Utah subsets
tapply(BIL$MW_PM_1, BIL$Utah, mean, na.rm=TRUE)

#differences of means test (one-tailed)
t.test(BIL$MW_PM_1~BIL$Utah, mu = 0, alternative = "less")

#standard deviation for NotUtah and Utah subsets
tapply(BIL$MW_PM_1, BIL$Utah, sd, na.rm=TRUE)

#differences of variation test (one-tailed)
var.test(BIL$MW_PM_1~BIL$Utah, ratio = 1, alternative = "greater")
```

Organizational commitment (Comm_1)
```{r}
#mean of whole sample
mean(BIL$Comm_1, na.rm=TRUE)

#standard deviation of whole sample
sd(BIL$Comm_1, na.rm=TRUE)

#mean for NotUtah and Utah subsets
tapply(BIL$Comm_1, BIL$Utah, mean, na.rm=TRUE)

#differences of means test (one-tailed)
t.test(BIL$Comm_1~BIL$Utah, mu = 0, alternative = "less")

#standard deviation for NotUtah and Utah subsets
tapply(BIL$Comm_1, BIL$Utah, sd, na.rm=TRUE)

#differences of variation test (one-tailed)
var.test(BIL$Comm_1~BIL$Utah, ratio = 1, alternative = "greater")
```

Student's Choice 1: Good healthcare (Healthcare)
```{r}
#mean of whole sample
mean(BIL$Healthcare, na.rm=TRUE)

#standard deviation of whole sample
sd(BIL$Healthcare, na.rm=TRUE)

#mean for NotUtah and Utah subsets
tapply(BIL$Healthcare, BIL$Utah, mean, na.rm=TRUE)

#differences of means test (one-tailed)
t.test(BIL$Healthcare~BIL$Utah, mu = 0, alternative = "less")

#standard deviation for NotUtah and Utah subsets
tapply(BIL$Healthcare, BIL$Utah, sd, na.rm=TRUE)

#differences of variation test (one-tailed)
var.test(BIL$Healthcare~BIL$Utah, ratio = 1, alternative = "greater")
```

Student's Choice 2: Respected at work (Respect)
```{r}
#mean of whole sample
mean(BIL$Respect, na.rm=TRUE)

#standard deviation of whole sample
sd(BIL$Respect, na.rm=TRUE)

#mean for NotUtah and Utah subsets
tapply(BIL$Respect, BIL$Utah, mean, na.rm=TRUE)

#differences of means test (one-tailed)
t.test(BIL$Respect~BIL$Utah, mu = 0, alternative = "less")

#standard deviation for NotUtah and Utah subsets
tapply(BIL$Respect, BIL$Utah, sd, na.rm=TRUE)

#differences of variation test (one-tailed)
var.test(BIL$Respect~BIL$Utah, ratio = 1, alternative = "greater")
```

Student's Choice 3: Empathetic leader (EmpatheticLeader)
```{r}
#mean of whole sample
mean(BIL$EmpatheticLeader, na.rm=TRUE)

#standard deviation of whole sample
sd(BIL$EmpatheticLeader, na.rm=TRUE)

#mean for NotUtah and Utah subsets
tapply(BIL$EmpatheticLeader, BIL$Utah, mean, na.rm=TRUE)

#differences of means test (one-tailed)
t.test(BIL$EmpatheticLeader~BIL$Utah, mu = 0, alternative = "less")

#standard deviation for NotUtah and Utah subsets
tapply(BIL$EmpatheticLeader, BIL$Utah, sd, na.rm=TRUE)

#differences of variation test (one-tailed)
var.test(BIL$EmpatheticLeader~BIL$Utah, ratio = 1, alternative = "greater")
```

Student's Choice 4: Turnover intentions (Turnover)
```{r}
#mean of whole sample
mean(BIL$Turnover, na.rm=TRUE)

#standard deviation of whole sample
sd(BIL$Turnover, na.rm=TRUE)

#mean for NotUtah and Utah subsets
tapply(BIL$Turnover, BIL$Utah, mean, na.rm=TRUE)

#differences of means test (one-tailed)
t.test(BIL$Turnover~BIL$Utah, mu = 0, alternative = "greater")

#standard deviation for NotUtah and Utah subsets
tapply(BIL$Turnover, BIL$Utah, sd, na.rm=TRUE)

#differences of variation test (one-tailed)
var.test(BIL$Turnover~BIL$Utah, ratio = 1, alternative = "greater")
```




**Proportions for Whole Sample**
```{r}
tab1(BIL$HighIncome, graph = FALSE)
tab1(BIL$CollegeDegree, graph = FALSE)
tab1(BIL$SAL, graph = FALSE)
tab1(BIL$OrgYearsMoreThanTen, graph = FALSE)
tab1(BIL$Married, graph = FALSE)
```



**Subsetting Female Respondents**
```{r}
#Create a subset of only female respondents
Female <- subset(BIL, Gender == 1)
table(Female$Utah)
```



**Proportions Female in Utah and Female not in Utah**
(prop.test is run as Female Utah first, Female Not Utah second, so a positive z means the Utah share is higher.
z is the square root of X-squared, given the sign of the difference.
All tests are one-tailed: "greater" tests Female Utah > Female Not Utah and "less" tests Female Utah < Female Not Utah.)

Income above $100,000 (HighIncome)
```{r}
FemaleUtahHighIncome <- table(Female$Utah, Female$HighIncome)
FemaleUtahHighIncome <- addmargins(FemaleUtahHighIncome)
FemaleUtahHighIncome
FemaleUtahHighIncome.test <- prop.test(x = c(FemaleUtahHighIncome["Utah","1"], FemaleUtahHighIncome["NotUtah","1"]),
                    n = c(FemaleUtahHighIncome["Utah","Sum"], FemaleUtahHighIncome["NotUtah","Sum"]),
                    correct = F, alternative = "less")
FemaleUtahHighIncome.test
#z statistic
sign(FemaleUtahHighIncome.test$estimate[1] - FemaleUtahHighIncome.test$estimate[2]) * sqrt(FemaleUtahHighIncome.test$statistic)
```

College Degree (CollegeDegree)
```{r}
FemaleUtahCollegeDegree <- table(Female$Utah, Female$CollegeDegree)
FemaleUtahCollegeDegree <- addmargins(FemaleUtahCollegeDegree)
FemaleUtahCollegeDegree
FemaleUtahCollegeDegree.test <- prop.test(x = c(FemaleUtahCollegeDegree["Utah","1"], FemaleUtahCollegeDegree["NotUtah","1"]),
                    n = c(FemaleUtahCollegeDegree["Utah","Sum"], FemaleUtahCollegeDegree["NotUtah","Sum"]),
                    correct = F, alternative = "greater")
FemaleUtahCollegeDegree.test
#z statistic
sign(FemaleUtahCollegeDegree.test$estimate[1] - FemaleUtahCollegeDegree.test$estimate[2]) * sqrt(FemaleUtahCollegeDegree.test$statistic)
```

In Sales or Business Development (SAL, 1 Yes, 2 No) (SAL)
```{r}
FemaleUtahSAL <- table(Female$Utah, Female$SAL)
FemaleUtahSAL <- addmargins(FemaleUtahSAL)
FemaleUtahSAL
FemaleUtahSAL.test <- prop.test(x = c(FemaleUtahSAL["Utah","1"], FemaleUtahSAL["NotUtah","1"]),
                    n = c(FemaleUtahSAL["Utah","Sum"], FemaleUtahSAL["NotUtah","Sum"]),
                    correct = F, alternative = "less")
FemaleUtahSAL.test
#z statistic
sign(FemaleUtahSAL.test$estimate[1] - FemaleUtahSAL.test$estimate[2]) * sqrt(FemaleUtahSAL.test$statistic)
```

More than 10 years in organization (OrgYearsMoreThanTen)
```{r}
FemaleUtahOrgYearsMoreThanTen <- table(Female$Utah, Female$OrgYearsMoreThanTen)
FemaleUtahOrgYearsMoreThanTen <- addmargins(FemaleUtahOrgYearsMoreThanTen)
FemaleUtahOrgYearsMoreThanTen
FemaleUtahOrgYearsMoreThanTen.test <- prop.test(x = c(FemaleUtahOrgYearsMoreThanTen["Utah","1"], FemaleUtahOrgYearsMoreThanTen["NotUtah","1"]),
                    n = c(FemaleUtahOrgYearsMoreThanTen["Utah","Sum"], FemaleUtahOrgYearsMoreThanTen["NotUtah","Sum"]),
                    correct = F, alternative = "less")
FemaleUtahOrgYearsMoreThanTen.test
#z statistic
sign(FemaleUtahOrgYearsMoreThanTen.test$estimate[1] - FemaleUtahOrgYearsMoreThanTen.test$estimate[2]) * sqrt(FemaleUtahOrgYearsMoreThanTen.test$statistic)
```

Student Choice: Married or cohabitating (Married)
```{r}
FemaleUtahMarried <- table(Female$Utah, Female$Married)
FemaleUtahMarried <- addmargins(FemaleUtahMarried)
FemaleUtahMarried
FemaleUtahMarried.test <- prop.test(x = c(FemaleUtahMarried["Utah","1"], FemaleUtahMarried["NotUtah","1"]),
                    n = c(FemaleUtahMarried["Utah","Sum"], FemaleUtahMarried["NotUtah","Sum"]),
                    correct = F, alternative = "greater")
FemaleUtahMarried.test
#z statistic
sign(FemaleUtahMarried.test$estimate[1] - FemaleUtahMarried.test$estimate[2]) * sqrt(FemaleUtahMarried.test$statistic)
```
