---
title: "R Notebook"
output: html_notebook
---

**MGMT 3345 Unit 1 Project**

```{r setup, include=FALSE}
# Packages needed for this project.
# install.packages(c("readxl")) # uncomment/run once if not already installed
library(readxl)
```

**STEP 1**
#Complete this step in Excel.

Real returns were calculated in Excel using the CPI-deflation method:

    Real Return (%) = ( (P_t / P_t-1) / (CPI_t / CPI_t-1) - 1 ) * 100

Per the project instructions, the CPI for October 2025 was set to the
Treasury Department's contingency index of 325.604 (BLS did not release an
official October 2025 CPI due to the government shutdown). See the
"Source Data" tab of `Fidelity_Real_Returns.xlsx` for the priced/CPI inputs
and the "Fidelity" tab for the resulting 58 real-return observations
(Feb 2021 - Nov 2025).

Import the completed real returns file for use in the remaining steps.

```{r import-data}
# Update this path if your file lives somewhere else on your machine.
df <- read_excel("data/Fidelity_Real_Returns.xlsx", sheet = "Fidelity")

# Row 1 (Jan 2021) has no prior month price, so it has no real return.
# Drop it, leaving the 58 monthly real-return observations (Feb 2021-Nov 2025).
df <- df[-1, ]

str(df)
```

**STEP 2**
#Descriptive statistics

```{r descriptive-stats}
funds <- c("Consumer_Staples", "Energy", "Financials",
           "Health_Care", "Information_Technology", "Utilities")

fund_mean <- sapply(df[funds], mean)
fund_sd   <- sapply(df[funds], sd)

descriptive_stats <- data.frame(
  Fund               = funds,
  `Monthly Return (Mean)` = round(fund_mean, 4),
  `Standard Deviation`    = round(fund_sd, 4),
  check.names = FALSE
)

descriptive_stats
```

Copy these mean and standard deviation values (rounded to 4 decimal
places) into the Step 2a table in the Word document.

**STEP 3**
#Graphs

```{r line-graphs}
plot(df$Consumer_Staples ~ df$Date, main = "Consumer Staples: Monthly Real Return",
     xlab = "Date", ylab = "Real Return (%)", col = "darkgreen", type = "l",
     ylim = c(-25, 25))

plot(df$Energy ~ df$Date, main = "Energy: Monthly Real Return",
     xlab = "Date", ylab = "Real Return (%)", col = "darkorange", type = "l",
     ylim = c(-25, 25))

plot(df$Financials ~ df$Date, main = "Financials: Monthly Real Return",
     xlab = "Date", ylab = "Real Return (%)", col = "steelblue", type = "l",
     ylim = c(-25, 25))

plot(df$Health_Care ~ df$Date, main = "Health Care: Monthly Real Return",
     xlab = "Date", ylab = "Real Return (%)", col = "purple", type = "l",
     ylim = c(-25, 25))

plot(df$Information_Technology ~ df$Date, main = "Information Technology: Monthly Real Return",
     xlab = "Date", ylab = "Real Return (%)", col = "red", type = "l",
     ylim = c(-25, 25))

plot(df$Utilities ~ df$Date, main = "Utilities: Monthly Real Return",
     xlab = "Date", ylab = "Real Return (%)", col = "black", type = "l",
     ylim = c(-25, 25))
```

**Step 3b. What can you say about the returns for each company?**

*(TODO - write your own response here, in complete sentences. Look at
the six graphs above and the mean/standard deviation table from Step 2:
which funds swing the most from month to month, which stay flat, do any
funds move together or in opposite directions at the same points in
time, etc. Comment on similarities and differences you observe.)*

**STEP 4**
#Portfolios

Each portfolio invests 40% in the first fund listed and 60% in the
second fund listed.

*(TODO - the three pairs below are placeholders so the code runs end to
end. Swap in the two funds you actually want for each portfolio, then
re-run this chunk and copy the results into the Word document. Update
the fund names in each `portfolio()` call and give each portfolio a
short label that matches what you write in the Word doc.)*

```{r portfolio-function}
portfolio <- function(x, y, w1 = 0.4, w2 = 0.6) {
  rho <- cor(x, y)
  exp_return <- w1 * mean(x) + w2 * mean(y)
  variance <- (w1^2) * var(x) + (w2^2) * var(y) + 2 * w1 * w2 * cov(x, y)
  std_dev <- sqrt(variance)

  round(c(
    Correlation      = rho,
    `Expected Return` = exp_return,
    Variance         = variance,
    `Standard Deviation` = std_dev
  ), 4)
}
```

```{r portfolio-1}
# Portfolio 1: Consumer_Staples (40%) / Utilities (60%)
portfolio(df$Consumer_Staples, df$Utilities)
```

```{r portfolio-2}
# Portfolio 2: Energy (40%) / Information_Technology (60%)
portfolio(df$Energy, df$Information_Technology)
```

```{r portfolio-3}
# Portfolio 3: Financials (40%) / Health_Care (60%)
portfolio(df$Financials, df$Health_Care)
```

Copy each portfolio's correlation coefficient, expected return, variance,
and standard deviation (rounded to 4 decimal places) into the Step 4
tables in the Word document, and state which two funds you chose for
each one.

**STEP 5**
#Analysis (complete in the Word document)

*(TODO - Steps 5a and 5b are written analysis and personal investment
judgment calls based on the Step 4 numbers above - not something to fill
in here. Use the expected return / standard deviation values you copied
into the Word document to support your recommendations.)*
